---
title: "Data wrangling for beetles biodiversity"
author: "Julia Barreto"
date: "8/23/2019"
output: 
  rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, message = F, warning = F, error= F)
library(here)
library(tidyverse)
library(kableExtra)
library(readxl)
library(janitor)
```

# Importing and organizing datasets:

Raw biodiversity data (beetles abundance):
```{r}
data <- read.csv(here("datasets", "beetles_biodiversity.csv"), header = TRUE,
                 sep=";")
```

## PREP AND TREATING DATA

### Remove singletons:

Check species of only one individual collected in the whole project to remove it. The detected singletons from forests are:
```{r, echo=T}
spp_all <- data %>% 
  dplyr::select(-Landscape, -Point, -Trap_num,-Year,
                -Month,-Day,-Muncipality,-Lat,-Long,-Alt)

(singletons <- data.frame(sort(colSums(spp_all))) %>% 
  filter(sort.colSums.spp_all..==1) %>% 
  rownames() )

### we have 6 project singletons

# remove those project singletons from data
data2 <- data %>% 
  dplyr::select(-all_of(singletons))
```
### Tidy format for analyses

For now we have species per column in the abundance matrix per pitfall, tidy formatting to remove columns we won't use and pivot the table longer to organize species abundance in rows instead of columns. Data product renamed as 'beetles'. 

```{r}
data3 <- data2 %>% 
  dplyr::select(-Trap_num,-Year,-Month,-Day,-Muncipality,-Lat,-Long,-Alt) %>%
  group_by(Landscape,Point) %>% summarise_all(sum) %>% gather("spp","abundance",3:52)

# rename data 5 so I can remember what it stands for
beetles <- data3
```

## HABITAT ASSOCIATION DATA

Dataset containing trait information of each species to classify into habitat association based on literature data and personal  data from  an expert taxonomist (Prof. Dr. Fernando Augusto B. Silva, coauthor)

```{r}
hab.data <- read.csv(here("datasets", "2022nov_spp_trait_revFB.csv"), 
                     sep= ";", header= T) %>%
  filter(Code %in% unique(beetles$spp)) %>%
  dplyr::select(-spp) %>%
  dplyr::rename("spp"= Code)
 
# # # of species 50
# length(unique(hab.data$spp))
# 
# # # of genus 17
# length(unique(hab.data$Genus))

# # spp per class
hab.data %>% tabyl(Classification)

# # # of individuals per class (need to run beta RC)
# b.com %>% ungroup() %>%
#   dplyr::select(all_of(fs.spp)) %>%
#   sum()
#   
# b.com %>% ungroup() %>%
#   dplyr::select( all_of(nfs.spp)) %>%
#   sum()


```


## FOREST COVER

Data on forest cover percentage of the 12 focal landscapes measured at a 3 and 5 km radius measured based in MapBiomas maps in June 2019. We used the Sampling Design tool in ArcGIS 10.1 to calculate the percentage of native forest cover in circular buffers of 3km and 5km radii around the landscape’s centroids.

```{r}
env.data <- read.csv(here("datasets", "perc_fc_JB.csv"), 
                     sep = ";", dec = ",") %>% 
              dplyr::rename(fc_3km= perfc_3km, fc_5km= perfc_5km) %>%
  dplyr::select(Landscape, fc_3km, fc_5km)
```

# CALCULATING METRICS

## ALPHA, BETA & GAMMA: FS & NFS

Using two datasets separated into habitat association classes: FS- forest specialists and NFS- non-forest specialists.
```{r}
alpha_hab <- beetles %>% mutate(occurrence= 1*(abundance>0)) %>%
  left_join(hab.data,by= "spp") %>%
  group_by(Landscape,Point, Classification) %>% 
  summarise(richness=sum(occurrence)) %>% # dois locais tem riqueza zero 263-13 e 266-28
  group_by(Landscape, Classification) %>% # now per landscape n=12
  summarise(alpha_hab = round(mean(richness),2)) %>%
  spread(Classification, alpha_hab) %>%
  rename(alpha_FS= FS, alpha_NFS= NFS)
  
  
## gamma richness have to consider spp not to count it twice as it appears more than once per landscape

gamma_hab <-beetles %>%  mutate(occurrence= 1*(abundance>0)) %>%
  left_join(hab.data,by= "spp") %>%
  group_by(Landscape,Classification, spp) %>% 
  summarise(tot.occ=sum(occurrence)) %>%
  filter(tot.occ !=0) %>% # removing spp that were blank in that landscape
  group_by(Landscape, Classification) %>%
  summarise(gamma_hab = n()) %>%
  spread(Classification, gamma_hab) %>%
  rename(gamma_FS= FS, gamma_NFS= NFS)

div_hab <- gamma_hab %>% left_join(alpha_hab,by= 'Landscape') %>% mutate(beta_FS= round(gamma_FS/alpha_FS,2), beta_NFS= round(gamma_NFS/alpha_NFS,2))

kable(div_hab, booktabs=T, row.names = F,caption= "Dataset to test alpha, gamma and beta diversity of Forest Specialists (FS) species and Non-Forest Specialist species.") %>% kable_styling(full_width = F)
```

# ABUNDANCES: FS & NFS
Get the beetles' species abundance matrix, calculate abundance at site and landscape level:

```{r}
st.hab_ab <- beetles %>%
  left_join(hab.data,by= "spp") %>%
  group_by(Classification) %>%
  mutate(ab.tot = sum(abundance)) %>%
  group_by(Landscape, Point, Classification, ab.tot) %>% 
  summarise(ab.point=sum(abundance)) %>% 
  group_by(Landscape, Classification, ab.tot) %>% 
  mutate(m.ab = round(mean(ab.point),2)) %>%
  gather(ab.spread, value, -(Landscape:Point:Classification)) %>%
  unite(temp, Classification, ab.spread) %>%
  spread(temp, value) %>%
  mutate(ab.land_FS= sum(FS_ab.point), ab.land_NFS= sum(NFS_ab.point))

kable(st.hab_ab) %>% kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)

ab_hab <- st.hab_ab %>% group_by(Landscape, ab.land_FS, ab.land_NFS) %>% 
  summarise(mean.ab_FS= round(mean(FS_ab.point), 2), 
            mean.ab_NFS= round(mean(NFS_ab.point), 2), 
            med.ab_FS= median(FS_ab.point), med.ab_NFS= median(NFS_ab.point), 
            cv.ab_FS= round(sd(FS_ab.point)/mean(FS_ab.point), 2), 
            cv.ab_NFS= round(sd(NFS_ab.point)/mean(NFS_ab.point), 2))

kable(ab_hab) %>% kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```
# BETA Raup-Crick: FS & NFS

Load functions:
```{r}
# Load beta RC presence-absence function:
source(here("scripts", "raup_crick_modif.R"))
# load beta RC abundance-based function:
# função rc_qrand modificado de Puttker et al. 2014
source(here("scripts", "rc_qrand_modif.R"))
```

Start from community matrix (spp x site) and subset it for each group, as we treat it separately:
```{r}
## community data (sppXsite matrix) - entry for beta RC calculations
b.com <- data2 %>% 
  dplyr::select(-Trap_num,-Year,-Month,-Day,-Muncipality,-Lat,-Long,-Alt) %>%
  group_by(Landscape,Point) %>% summarise_all(sum)

## vector for each set of species
fs.spp <- hab.data %>% filter(Classification=="FS") %>% pull(spp)
nfs.spp <-  hab.data %>% filter(Classification=="NFS") %>% pull(spp)

# sitXsp datasets for each
b.com.fs <- b.com %>% dplyr::select(Landscape, Point, all_of(fs.spp))
b.com.nfs <- b.com %>% dplyr::select(Landscape, Point, all_of(nfs.spp))
```


## Presence-absence beta RC for FS and NFS:
Set a regional pool as reference:
```{r}
## extract landscapes to loop it to
Landscape <- unique(b.com$Landscape)
##### FS #####
# Pool regional - Forest specialists that were caught in all lands
reg.pool.fs <- colSums(b.com.fs[,3:dim(b.com.fs)[2]]) ## regional pool of FS spp
rc.df.fs <- data.frame(Landscape=NA, beta.rc=NA)
beta.lands.fs <- list()

for (i in 1:length(Landscape)){
  b.com.fs2 <- b.com.fs[b.com.fs$Landscape == Landscape[i],]
  beta <- raup_crick(b.com.fs2[,3:dim(b.com.fs2)[2]], plot_names_in_col1 = F, occur = reg.pool.fs)
  rc.df.fs[i,1] <- Landscape[i]
  rc.df.fs[i,2] <- mean(beta)
  beta.lands.fs[[i]] <- beta
}

(rc.df.fs <- rc.df.fs %>% rename(brc.fs.raw = beta.rc))
```

Remove NA that arise from zero occurrence sites:
```{r}
# rmNA: remove 0 from calculations
ttes <- map(beta.lands.fs, as.matrix)
# function that substitute 0 per NA:
fufa2 <- function(x)  ifelse(x==0, NA, x) 
ttes <- modify_depth(ttes, 1, fufa2)
ttes <- modify_depth(ttes, 1, as.dist)

rc.df.fs$brc.fsNA <- map_dbl(ttes, mean, na.rm=T)
```

Same steps for non-forest specialists:
```{r}
##### NFS #####
reg.pool.nfs <- colSums(b.com.nfs[,3:dim(b.com.nfs)[2]]) ## regional pool of nfs spp
rc.df.nfs <- data.frame(Landscape=NA, beta.rc=NA)
beta.lands.nfs <- list()

for (i in 1:length(Landscape)){
  b.com.nfs2 <- b.com.nfs[b.com.nfs$Landscape == Landscape[i],]
  beta <- raup_crick(b.com.nfs2[,3:dim(b.com.nfs2)[2]], plot_names_in_col1 = F, occur = reg.pool.nfs)
  rc.df.nfs[i,1] <- Landscape[i]
  rc.df.nfs[i,2] <- mean(beta)
  beta.lands.nfs[[i]] <- beta
}

(rc.df.nfs <- rc.df.nfs %>% rename(brc.nfs.raw = beta.rc))

# rmNA: remove 0 from calculations
ttes <- map(beta.lands.nfs, as.matrix)
ttes <- modify_depth(ttes, 1, fufa2) # function that sub 0 for NA
ttes <- modify_depth(ttes, 1, as.dist)
rc.df.nfs$brc.nfsNA <- map_dbl(ttes, mean, na.rm=T)
```

## Abundance-based Beta RC for FS and NFS:

First for Forest specialists, then for non-forest specialists, we set a regional species group, caculate beta and transform zero occurrence sites in NA to remove it from calculations.

Final Beta Raup-Crick, both presence-absence abundance-based, below:
```{r}

##### RAUP-CRICK ABUNDANCE modif Puttker #####
#### FS ####

## set regional spp pool
reg.pool.fs <- b.com.fs %>% ungroup() %>% dplyr::select(-Landscape, -Point) %>% 
  summarise_all(sum) %>% gather("spp","abund",1:28)
reg.pool.fs <- as.data.frame(reg.pool.fs)

rc.df.abund.fs <- data.frame(Landscape=NA, brc.abund.fs.raw=NA)
beta.lands.abund.fs <- list()

for (i in 1:length(Landscape)){
  b.com.fs2 <- b.com.fs[b.com.fs$Landscape == Landscape[i],]
  b.com.fs2 <- as.matrix(b.com.fs2)
  beta <- rc_qrand(b.com.fs2[,3:dim(b.com.fs2)[2]], pool = reg.pool.fs)
  rc.df.abund.fs[i,1] <- Landscape[i]
  rc.df.abund.fs[i,2] <- mean(beta, na.rm = T)
  beta.lands.abund.fs[[i]] <- beta
}

# rmNA: remove 0 from calculations
ttes <- map(beta.lands.abund.fs, as.matrix)
ttes <- modify_depth(ttes, 1, fufa2) # function that sub 0 per NA
ttes <- modify_depth(ttes, 1, as.dist)
rc.df.abund.fs$brc.abund.fsNA <- map_dbl(ttes, mean, na.rm= T)

#### NFS ####

## set regional spp pool
reg.pool.nfs <- b.com.nfs %>% ungroup() %>% 
  dplyr::select(-Landscape, -Point) %>% 
  summarise_all(sum) %>% gather("spp","abund",1:22)
reg.pool.nfs <- as.data.frame(reg.pool.nfs)

rc.df.abund.nfs <- data.frame(Landscape=NA, brc.abund.nfs.raw=NA)
beta.lands.abund.nfs <- list()

for (i in 1:length(Landscape)){
  b.com.nfs2 <- b.com.nfs[b.com.nfs$Landscape == Landscape[i],]
  beta <- rc_qrand(b.com.nfs2[,3:dim(b.com.nfs2)[2]], pool = reg.pool.nfs)
  rc.df.abund.nfs[i,1] <- Landscape[i]
  rc.df.abund.nfs[i,2] <- mean(beta, na.rm = T)
  beta.lands.abund.nfs[[i]] <- beta
}

# rmNA: remove 0 from calculations
ttes <- map(beta.lands.abund.nfs, as.matrix)
ttes <- modify_depth(ttes, 1, fufa2) # function that sub 0 per NA
ttes <- modify_depth(ttes, 1, as.dist)
rc.df.abund.nfs$brc.abund.nfsNA <- map_dbl(ttes, mean, na.rm= T)

#### gather all data in one set to then save it
(rc.df <- rc.df.fs %>%  # P/A-based Brc  for fs spp
   left_join(rc.df.nfs, by= "Landscape") %>%# P/A-based Brc for NFS
   left_join(rc.df.abund.fs, by= "Landscape") %>% # Abund-based FS
   left_join(rc.df.abund.nfs, by= "Landscape") %>% # Abund-based NFS
   left_join(env.data, by= "Landscape"))
```

# Save final tidy dataset:
```{r}
save(beetles, # raw data for species occurrence per landscape/site; 
     hab.data, # data containing classes of habitat association;
     env.data, # environmental variables
     div_hab, # diversity (alpha, beta and gamma) between classification habitat association
     ab_hab, # mean and median abundance between classification habitat association
     st.hab_ab, # abundance data between classification habitat association per site
     rc.df, # Beta RC calculations
     file= here("datasets", "beetles.Rdata"))
```